\mytitle

### 2020-04-14 paper.tex
Edited Abstract and Introduction.

\begin{abstract} 
%Wind farms play an increasingly important
%role as a source of renewable energy and 
The wind farm layout optimization (WFLO) problem concerns
selecting turbine positions to maximize energy production. While a number of heuristic and exact techniques have been proposed in the literature to solve WFLO, it is natural to model the interaction between turbines as a quadratic objective function as, indeed, has been done in some existing optimization models. Recent advances in optimization hardware and software have both targeted quadratic constraints: commercial mixed integer linear solvers have been extended to quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving Quadratic Unconstrained Binary Optimization (QUBO) problems. 
%It is a key problem 
%when designing a new wind farm, as it has a direct influence
%on the amount of the total produced energy. 
%In recent years there have been advances in both software technologies (e.g., the extension of commercial mixed integer linear programming solvers to handle quadratic problems) and nascent hardware development (e.g., quantum computing chips and specialized classical chips) that have relevance for the WFLO problem. In both cases, the advances have 
%Recent years saw 
%advances in software and hardware optimization technologies, such as Gurobi's quadratic solver 
%and Fujitsu's digital annealer (DA), respectively. 
%These technologies create an opportunity to improve 
%over current wind farm optimization solutions
%both in terms of solution quality and solve time. 
%In order to exploit these technologies, 
%we must represent the WFLO problem as 
%a quadratic optimization model. Specifically, in order 
%to apply a software-based optimization
%solution (Gurobi), 
In this paper, we 
introduce two novel quadratic programming models for WFLO: a quadratic constrained optimization problem (QCOP) with binary decision variables and a QUBO.
%Subsequently, we provide the unconstrained counterpart of the QCOP, 
%namely the quadratic unconstrained binary optimization (QUBO), which  
%is the underlying representation for quantum-inspired 
%optimization hardware such as the DA. 
A thorough empirical 
evaluation using a commercial solver and a specialized classical chip for QUBO solving demonstrates the strengths of the two models. 
Our results show that both approaches yield fast and high-quality solutions 
and
improve over existing WFLO solutions. In particular, the QUBO model excels at finding high quality solutions very quickly while the QCOP model is able to find better solutions and provide quality guarantees over a longer run-time.
\end{abstract}
\begin{keyword}
	Wind Farm Layout Optimization (WFLO) \sep Quadratic Programming (QP) \sep Quadratic Unconstrained Binary Optimization (QUBO) \sep Digital Annealer
\end{keyword}

@@

\section{Introduction}

We revisit 
the wind farm layout optimization (WFLO) problem 
that was first considered 
in~\citet{MOSETTI1994105}. 
In WFLO, we aim to place wind turbines 
in a predefined area to maximize 
the total generated power. 
Turbine location decisions    
influence
the
total power production due to  
interference effects, aka wakes, generated by upstream 
turbines. Furthermore, 
there are proximity constraints between turbines
that prevent from turbines to be placed next to each other.
These inter-turbine effects 
are naturally captured using quadratic constraints and objective
functions. Several exact optimization approaches have 
been proposed to solve
the WFLO problem~\cite{turner2014new,Zhang2014}. However, these
approaches represented the quadratic effects and constraints
using linear models that approximate the true dependencies.

In this work, we introduce two 
novel quadratic WFLO models. The first model 
represents WFLO using a 
quadratic constrained optimization problem (QCOP). 
The QCOP can be solved using state-of-the-art solvers that
specialize in such problems, e.g., Gurobi~\cite{gurobi}. 
Our second model is based on a quadratic unconstrained binary optimization (QUBO) representation of the WFLO 
problem.
This enables the use of specialized optimization hardware tailored to solve
QUBOs. In this paper, we use Fujitsu's digital annealer, a simulated annealing based QUBO solver, to demonstrate the strength
of our approach. The main contributions of this paper are: 
\begin{enumerate} 
\item We propose a novel quadratic modeling framework for WFLO that captures wake 
effects directly through their linear superposition 
without the need for approximations and ad-hoc solutions,
\item We provide the computational complexity of WFLO and show that it is $\mathcal{NP}$-hard, and, 
\item We test our quadratic formulations
on a commerical software-based quadratic solver (Gurobi) and on Fujitsu's digital 
annealer. 
\end{enumerate} Our experiments are focused on the performance of the two approaches
compared to 
existing optimization methods. To this end,
we experiment with 12 standard WFLO benchmark problems (c.f.~\cite{turner2014new})
and show that solving quadratic WFLO models
using Gurobi and the DA
achieves state-of-the-art performance, often outperforming
existing solutions when provided sufficient run-time. Furthermore, we
we observe that the digital 
annealer often provides order of magnitude faster solutions
compared to software-based approaches.

>>>>

\begin{abstract} 
%Wind farms play an increasingly important
%role as a source of renewable energy and 
The wind farm layout optimization (WFLO) problem requires the
selection of wind turbine positions to maximize energy production. While a number of heuristic and exact techniques have been proposed in the literature to solve WFLO, it is natural to model the interaction between turbines as a quadratic objective function as, indeed, has been done in some existing optimization models. Recent advances in optimization hardware and software have both targeted quadratic constraints: commercial mixed integer linear solvers have been extended to address some quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving Quadratic Unconstrained Binary Optimization (QUBO) problems. 
%It is a key problem 
%when designing a new wind farm, as it has a direct influence
%on the amount of the total produced energy. 
%In recent years there have been advances in both software technologies (e.g., the extension of commercial mixed integer linear programming solvers to handle quadratic problems) and nascent hardware development (e.g., quantum computing chips and specialized classical chips) that have relevance for the WFLO problem. In both cases, the advances have 
%Recent years saw 
%advances in software and hardware optimization technologies, such as Gurobi's quadratic solver 
%and Fujitsu's digital annealer (DA), respectively. 
%These technologies create an opportunity to improve 
%over current wind farm optimization solutions
%both in terms of solution quality and solve time. 
%In order to exploit these technologies, 
%we must represent the WFLO problem as 
%a quadratic optimization model. Specifically, in order 
%to apply a software-based optimization
%solution (Gurobi), 
In this paper, we 
introduce two novel quadratic programming models for WFLO: a quadratic constrained optimization problem (QCOP) with binary decision variables and a QUBO.
%Subsequently, we provide the unconstrained counterpart of the QCOP, 
%namely the quadratic unconstrained binary optimization (QUBO), which  
%is the underlying representation for quantum-inspired 
%optimization hardware such as the DA. 
A thorough empirical 
evaluation using a commercial solver and a specialized classical chip for QUBO solving demonstrates the strengths of the two models. 
Our results show that both approaches yield fast and high-quality solutions 
and
improve over existing WFLO solutions. In particular, the QUBO model delivers high quality solutions very quickly while the QCOP model can be used to find better solutions and provide quality guarantees over a longer run-time.
\end{abstract}
\begin{keyword}
	Wind Farm Layout Optimization (WFLO) \sep Quadratic Programming (QP) \sep Quadratic Unconstrained Binary Optimization (QUBO) \sep Digital Annealer
\end{keyword}

@@

\section{Introduction}
% Todo
The goal in the wind farm layout optimization (WFLO)
problem~\cite{MOSETTI1994105} is to locate wind turbines within a
predefined area to maximize the total power harvested from the wind
stream.  While physical considerations based on the diameter of the
blades prevent turbines being placed too close to each other, turbine
location decisions influence the total power production due to
interference effects (i.e., wakes) generated by upstream turbines.
These inter-turbine effects are naturally captured using quadratic
constraints and objective functions. However, with a few exceptions (e.g., \cite{turner2014new}), exact optimization
approaches that have been proposed for the WFLO
problem~\cite{Zhang2014,kuo2016wind} \todo{[Add more references here]}
represent the quadratic relations using linear models
that approximate the true dependencies.

In this work, we introduce two novel quadratic WFLO models. The first
model represents WFLO with a quadratic constrained optimization
problem (QCOP).  The QCOP can be solved using state-of-the-art solvers
that specialize in such problems, e.g., Gurobi~\cite{gurobi}.  Our
second model is based on a quadratic unconstrained binary optimization
(QUBO) representation of the WFLO problem.  Such a formulation enables the use of
specialized optimization hardware tailored to solve QUBOs. In this
paper, we use Fujitsu's Digital Annealer (DA), a hardware-based
QUBO solver, to demonstrate the strength of our approach. The main
contributions of this paper are:
\begin{enumerate} 
\item We propose a novel quadratic modeling framework for WFLO that captures wake 
effects directly through their linear superposition. 
%without the need for approximations and ad-hoc solutions,
\item We prove that WFLO is $\mathcal{NP}$-hard.
\item Through numerical experiments on two instantiations of our quadratic framework using Fujitsu's Digital Annealer and a commercial software-based quadratic solver (Gurobi), we demonstrate state-of-the-art performance in, respectively, quickly finding high-quality solutions and finding and proving optimal solutions over a longer run-time.
\end{enumerate}

Our experiments are focused on the performance of the two approaches
compared to 
existing optimization methods. To this end,
we experiment with 12 standard WFLO benchmark problems (c.f.~\cite{turner2014new})
and show that solving quadratic WFLO models
using Gurobi and the DA
achieves state-of-the-art performance, often outperforming
existing solutions when provided sufficient run-time. Furthermore, we
we observe that the DA 
often provides solutions of similar quality an order of magnitude faster
than software-based approaches.
